\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion} \markboth{CONCLUSION}{}

We have, as proposed in the paper \cite{REV}, implemented an interpreter for each of the reversible flowchart languages RL and SRL. We have described, and argued for, a series of changes to the syntax and feature set of the languages that we have been able to implement successfully. Furthermore, a program inverter and a program translator have been written for each language. Two interfaces, used to run the interpreters and perform the program transformations, have been designed; a command-line interface and a web-based user interface, both supporting all implemented features. The behaviours of the interpreters, the inverters, and the translators have been tested thoroughly, and the results have strongly indicated that everything indeed works as intended.

Potential bottlenecks of the implementation have been identified and discussed; most notable are the representations of programs, the efficiency of the interpreters, and the growth in size of programs being translated. Furthermore, we have proposed some ideas that could solve these problems if implemented. Further development is most certainly justified and --- perhaps --- even recommended.

We have written various programs in RL and SRL to demonstrate the usefulness of the languages; some are trivial, demonstrating one single feature, and some are non-trivial, demonstrating the languages as a whole to compute interesting functions reversibly.

%(Janus: \cite{JAN})

Thus, we find it safe to say that we have reached the goal of this project in a satisfactory way. % by implementing interpreters with correct behaviour and interfaces for interacting with these.
