
% går så videre til det mere specifikke med RL og SRL
% og snakker om den formelle syntaks
\section{RL and SRL}\label{sec:rlandsrl}
\noindent Reversible Language and Structured Reversible Language --- RL and SRL for short, repectively --- are two reversible flowchart languages proposed in~\cite{REV}. RL is an assembly-like language with unstructured jumps, while SRL, as the name indicates, is a similar language but with control structures and no jumps. The two languages share a common subset of features --- namely their expressions and step operations. The syntax of the two languages along with the common structures are given in~\cite[p.~99]{REV} and can be seen in Fig.~\ref{fig:rl_srl_syntax_and_structure}.\\

\noindent As we see, RL, described in Fig.~\ref{fig:rlspec}, consists of one or more (RL-)blocks, where each block is (uniquely) defined by a label and has a come-from assertion, zero or more step operations and an unstructured jump to any block in the program. Come-from assertions can either be conditional, unconditional, or an entry point. Similarly, jumps can either be conditional, unconditional, or an exit point. Note that a well-formed RL program must contain exactly one entry and one exit.

SRL is more linear; as we see in Fig.~\ref{fig:srlspec}, a program consists of exactly one (SRL-)block, where a block may either be an atomic step operation, a control structure --- that is, an \texttt{if}- or a \texttt{do-until} structure --- or recursively a sequence of two blocks. Sequences of blocks are executed in order, and thus, an SRL program looks a lot like a program written in common programming languages such as Python or C, to name a few. Note that a block in RL is different from a block in SRL.\\

\noindent Any atomic construct of each of these languages can be built using the building blocks of reversible flowcharts, and therefore it is clear that these are indeed reversible flowchart languages.\\

\noindent As mentioned, the two languages share expressions and step operations along with the basic value types. A brief overview follows:\\

\begin{description}
  \item[Values and types]~\
  \begin{itemize}
    \item A variable can either be mapped to an integer or to a list of integers.
    \item We only have integer literals; to obtain a certain list of integers, one has to bind each of the wanted values to a variable and push it to a list.
    \item Boolean values are represented by integers. A non-zero integer value will evaluate to true, and a zero integer value will evaluate to false.
  \end{itemize}
  \item [Step operations]~\
  \begin{itemize}
    \item To modify a variable, we have the variable update. The variable update only supports reversible operators, namely addition and subtraction. Indexing is supported in a single dimension. One rule is that the identifier being updated must not occur in the right-hand expression since such an update is irreversible.
    \item You can push the value of a variable onto a list with the \texttt{push} step operation. The variable being pushed must be zero-cleared afterwards.
    \item You can pop the top value from a list into a variable with the \texttt{pop} step operation. The variable must be zero-cleared beforehand.
    \item The \texttt{skip} step operation does nothing.
  \end{itemize}
  \item [Expressions]~\
  \begin{itemize}
    \item We can have any arbitrary binary operator between two expressions, recursively.
    \item An expression terminates with an integer literal or a variable lookup.
    \item The \texttt{top} expression is unary and returns the top of the operand.
    \item The \texttt{empty} predicate is unary and returns true if the operand is an empty list and false otherwise.
  \end{itemize}
\end{description}

% figur der viser grammatikken her tak
\begin{figure}[]

  \begin{subfigure}{\textwidth}
    \center
    $$\begin{matrix*}[l]
      {p} & ::= & {b} & \quad & {b} & ::= & a             & | & \texttt{ if }e\texttt{ then }b\texttt{ else }b\texttt{ fi }e\\
                 &     &           &        &            &  |  & b\ b & | & \texttt{ from }e\texttt{ do }b\texttt{ loop }b\texttt{ until }e\\
    \end{matrix*}$$
    \caption{Structured reversible language \textit{SRL}.}
    \label{fig:srlspec}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \center
    $$\begin{matrix*}[l]
      {q} & ::= & {d}^+                                      & & {k} & ::= & \texttt{from }l & & {j} & ::= & \texttt{goto }l\\
      {d} & ::= & {l}:\ {k}\ {a}^*\ {j} &       &            &  |  & \texttt{fi }e\texttt{ from }l\texttt{ else }l &
        & & | & \texttt{if }e\texttt{ goto }l\texttt{ else }l\\
    \end{matrix*}$$
    \caption{Unstructured reversible language \textit{RL}.}
    \label{fig:rlspec}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \center
    $$\begin{matrix*}[l]
      {a} & ::= & {x}\mathrel{\oplus}= e & \quad\quad\quad &  e & ::= & {c} \ | \ {x} \ | \ {x}[ e] \ | \  e\otimes e \ | \ \texttt{top }{x} \ | \ \texttt{empty }{x}\\
                 &  |  & {x}[ e]\mathrel{\oplus}= e &     & {c} & ::= & 0\ | \ 1\ |\ \cdots\ |\ 4294967295\\
                 &  |  & \texttt{push }{x}\ {x}      &     & \otimes    & ::= & \oplus \ | \ * \ | \ / \ | \ \cdots\\
                 &  |  & \texttt{pop  }{x}\ {x}      &     & \oplus     & ::= & +  \ | \  -  \ | \  \textasciicircum\\
                 &  |  & \texttt{skip}\\
    \end{matrix*}$$
    \caption{Reversible step operations and expressions.}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \center
    $$\begin{matrix*}[l]
      \mathtt{SRL}:     & {p}\in\mathtt{SRL}  & & {b}\in\mathtt{Blk}\\
      \mathtt{RL}:      & {q}\in\mathtt{RL}   & & {d}\in\mathtt{RLBlk} & & {j}\in\mathtt{Jump}  & & {k}\in\mathtt{From} & & {l}\in\mathtt{Label}\\
      \mathtt{SRL, RL}: & {a}\in\mathtt{Step} & &  e\in\mathtt{Exp}   & & {c}\in\mathtt{Const} & & {x}\in\mathtt{Var}  & & \oplus,\otimes\in\mathtt{Op}\\
    \end{matrix*}$$
    \caption{Syntax domains of \textit{SRL} and \textit{RL}.}\label{fig:common}
  \end{subfigure}

  \caption{Syntax of the two reversible flowchart languages.}
  \label{fig:rl_srl_syntax_and_structure}

\end{figure}
